---
title: "LSP Bayesian Regression on Linearized Double-logistic LSP"
author: "Matthew Shisler"
date: '2022-05-19'
output: html_document
---

```{r}
library(tidyverse)
source("utility_functions.R")
```

We will attempt to understand that effects of applying a Bayesian linear regression on the linearization (wrt to the parameter vector, \(\theta\)) of the double logistic function.

This simulation will involve different levels for three factors.
1) The magnitude of variation for each element of \(\theta\). (3 levels)
2) The magnitude of the measurement noise (i.e. error variance), \(\sigma^2\). (2 levels)
3) The number of observations per year. (3 levels)

We simulate 1000 evi2 curves for each combination of factor levels. 

We aggregate these 1000 evi2 curves and fit a frequentist non-linear regression to construct an "average model" estimate of the parameter vector, \(\theta_0\).

We linearize the double logistic function wrt to \(\theta\) centered on the \(\theta\) estimate from the average model. This constructs the basis function "data matrix."

For each simulated evi2 curve at a given combination of factor levels we fit a Bayesian linear regression to the linearized double-logistic function where the parameters are basis function coefficients \(b=(\theta-\theta_0)\). We assume an idealized case where we know the error variance and covariance matrix for \(\theta\) (the magnitudes of which are treated as different factor levels).

Finally, we construct 95% credible intervals for each evi2 curve using the posterior distribution for the Bayesian linear regression.

For results, we compute the coverage and widths of the 95% credible intervals and the the deviations of the \(\theta\) estimates from their true value. (estimated value - true value).

```{r}
sim_theta_mu  <- c(-1.80, 0.75, 120, 8, 290, 8, -7.750)

npar    <- length(sim_theta_mu)
sigma   <- c(0.025, 0.05)
nyear   <- 1000
nobs <- c(5,15,30)

sim_theta_var <- matrix(c( 0.01, 1e-5,  10, 1,  10, 1,  0.05,
                           0.04, 1e-4,  25, 2,  25, 2,  0.1,
                           0.08, 1e-3,  60, 4,  60, 4,  0.2),
                        nrow=3, ncol=npar, byrow=TRUE)

total_combos <- length(sigma)*length(nobs)*nrow(sim_theta_var)
coverages    <- data.frame(matrix(NA, nrow=total_combos, ncol=npar+3))
names(coverages) <- c(paste0("theta", 1:npar), "noise_level", "nobs_level", "theta_var_level")

sos_ci_widths <- matrix(NA, nrow=nyear,ncol=total_combos)
eos_ci_widths <- matrix(NA, nrow=nyear,ncol=total_combos)

sos_devs <- matrix(NA, nrow=nyear,ncol=total_combos)
eos_devs <- matrix(NA, nrow=nyear,ncol=total_combos)

low_high     <- c("low", "high")
low_med_high <- c("low", "med", "high")

index = 1
for(k in 1:2){
  for(j in 1:3){
    for(i in 1:3){
      # simulate the data
      sim <- sim_evi2(sim_theta_mu, sim_theta_var[i,], 
                      sigma=sigma[k], nyear=nyear,
                      min_obs = nobs[j], max_obs = nobs[j])
      
      # plot a fraction of the simualted datasets
      plot_evi2(sim, "parallel", frac=0.1)
      text(0,1, paste("noise level:", low_high[k]), pos=4)
      text(0,0.9, paste("num obs per yr:", nobs[j]), pos=4)
      text(0,0.8, paste("theta var level:", low_med_high[i]), pos=4)
      
      # create aggregated data set from % sample of simulated years
      sampled_years <- sample(nyear,floor(1*nyear))
      sim_agg <- do.call(rbind, sim$data[sampled_years])
      
      # construct average model by combining data over years (frequentist non-lin)
      # uses the nlsLM function of the minpack.lm package
      avg_model_over_years <- avg_fit(sim_agg = sim_agg, npar = npar)
      avg_model_theta <- coef(avg_model_over_years)
      
      # construct the basis functions from the 1st-order Taylor approx
      B <- basis_functions(1:365, avg_model_theta)
      # matplot(B, type = "l")
      # matplot(scale(B), type = "l")
      
      # priors 
      # In this idealized case we assume to know the residual 
      # variance and theta covariance.
      V   <- diag(sim_theta_var[i,])
      Q   <- solve(V)
      tau <- 1/sigma[k]^2
      
      # store results
      theta_hat <- theta_hat_lower <- theta_hat_upper <- sim$theta
      
      for (year in 1:nyear){
        
        # construct response and covariate matrix for bayesian regression
        DL_m  <- double_logistic(floor(sim$data[[year]]$t), sim_theta_mu)
        X     <- B[floor(sim$data[[year]]$t),]
        y     <- sim$data[[year]]$y_noise1-DL_m
      
        # compute posterior mean and covariance
        V <- solve(tau*t(X)%*%X + Q)
        M <- tau*V%*%t(X)%*%(y)
      
        # estimate theta and 95% credible interval
        theta_hat[year,]       <- avg_model_theta + as.numeric(M)
        theta_hat_lower[year,] <- qnorm(.025, mean=theta_hat[year,], sd=sqrt(diag(V)))
        theta_hat_upper[year,] <- qnorm(.975, mean=theta_hat[year,], sd=sqrt(diag(V)))
      }
      
      # check if true theta lies within the credible interval
      coverage <- theta_hat_lower<sim$theta & theta_hat_upper>sim$theta
      coverages[index,] <- c(colSums(coverage)/nyear, low_high[k], nobs[j], low_med_high[i])
      
      # record the widths of the CIs
      sos_ci_widths[,index] <- theta_hat_upper[,3]-theta_hat_lower[,3]
      eos_ci_widths[,index] <- theta_hat_upper[,5]-theta_hat_lower[,5]
      
      #record the estimated-true value deviation of sos and eos
      sos_devs[,index] <- theta_hat[,3] - sim$theta[,3]
      eos_devs[,index] <- theta_hat[,5] - sim$theta[,5]
      index=index+1
    }
  }
}
```

### Results

```{r}
coverages[,c(1:4,8:10)]
coverages[,c(5:7,8:10)]

sos_ci_widths <- reformat_sim_data(sos_ci_widths)
eos_ci_widths <- reformat_sim_data(eos_ci_widths)

ggplot(data=sos_ci_widths) +
  geom_boxplot(aes(x=num_obs, y=value, fill=theta_var_level)) +
  facet_grid(~noise_level, labeller = quick_facet_labeller) +
  labs(title = "Start-of-Season (theta3) CI Widths") +
  ylab("Days") +
  xlab("Number of Observations per Year") +
  theme_bw()

ggplot(data=eos_ci_widths) +
  geom_boxplot(aes(x=num_obs, y=value, fill=theta_var_level)) +
  facet_grid(~noise_level, labeller = quick_facet_labeller) +
  labs(title = "End-of-Season (theta5) CI Widths") +
  ylab("Days") +
  xlab("Number of Observations per Year") +
  theme_bw()

sos_devs <- reformat_sim_data(sos_devs)
eos_devs <- reformat_sim_data(eos_devs)

ggplot(data=sos_devs) +
  geom_boxplot(aes(x=num_obs, y=value, fill=theta_var_level)) +
  facet_grid(~noise_level, labeller = quick_facet_labeller) +
  labs(title = "Start-of-Season (theta3) Deviations (Estimate - True)") +
  ylab("Days") +
  xlab("Number of Observations per Year") +
  theme_bw()

ggplot(data=eos_devs) +
  geom_boxplot(aes(x=num_obs, y=value, fill=theta_var_level)) +
  facet_grid(~noise_level, labeller = quick_facet_labeller) +
  labs(title = "End-of-Season (theta5) Deviations (Estimate - True)") +
  ylab("Days") +
  xlab("Number of Observations per Year") +
  theme_bw()

```


